---
title: "AYU - Individual Week 2"
format: 
  html:
    toc: true
editor: visual
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
```

*Watch the videos to see how the similar problems are solve.*
---

### Parameter Estimators

**Problem 1** 

You are given the following data

| $y$| $x_1$   | $x_2$   |
|----|---|---|
| 2  | 1 | 1 |
| 3  | 1 | 1 |
| 4  | 2 | 2 |
| 6  | 3 | 2 |
| 10 | 3 | 4 |

You are using the following model:

$$
y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \epsilon_i, \quad i = 1, 2,..., 6
$$

You have determine 

$$
(X'X)^{-1} = \begin{bmatrix}
1.2 & -0.5 & 0\\
-0.5 & 0.75 & -0.5 \\
0 & -0.5 & 0.5
\end{bmatrix}
$$

Determine $\hat{\beta_1}$.

(A) -2.00
(B) -0.75
(C) -0.5
(D) 0.5
(E) 0.75


```{r}
X = matrix(c(1,1,1,1,1,1,1,2,3,3,1,1,2,2,4), ncol = 3)
y = matrix(c(2,3,4,6,10), ncol = 1)
solve(t(X)%*%X)
beta = solve(t(X)%*%X)%*%t(X)%*%y
beta
```

---

**Problem 2** 

You are given the following data

| $y$| $x_1$   | $x_2$   |
|----|---|---|
| 2  | 1 | 1 |
| 3  | 1 | 0 |
| 4  | 2 | 0 |
| 6  | 0 | 2 |
| 10 | 0 | 0 |

You are using the following model:

$$
y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \epsilon_i, \quad i = 1, 2,..., 6
$$

You have determine 

$$
(X'X)^{-1} = \begin{bmatrix}
0.8285714 & -0.4857143 & -0.4\\
-0.4857143  & 0.4571429 & 0.2 \\
-0.4000000  & 0.2000000 & 0.4
\end{bmatrix}
$$

Determine $\hat{\beta_0}$.

(A) -15.00
(B) -10.75
(C) 8.8
(D) 10.5
(E) 9.75


```{r, eval=FALSE}
X = matrix(c(1,1,1,1,1,1,1,2,0,0,1,0,0,2,0), ncol = 3)
y = matrix(c(2,3,4,6,10), ncol = 1)
solve(t(X)%*%X)
beta = solve(t(X)%*%X)%*%t(X)%*%y
beta
```

---

### F-test 

**Problem 3** (SRM - Sample Question 24)

Sarah performs a regression of the return on a mutual fund ($y$) on four predictors
plus an intercept. She uses monthly returns over 105 months.
Her software calculates the $F$ statistic for the regression as $F = 20.0$, but then it quits
working before it calculates the value of $R^2$. While she waits on hold with the help desk,
she tries to calculate $R^2$ from the F-statistic.

Determine which of the following statements about the attempted calculation is true.

(A) There is insufficient information, but it could be calculated if she had the value of
the residual sum of squares (RSS).
(B) There is insufficient information, but it could be calculated if she had the value of
the total sum of squares (TSS) and RSS.
(C) $R^2 = 0.44$
(D) $R^2 = 0.56$
(E) $R^2 = 0.80$


```{r}
n = 105
p = 4
F = 20

R2 = (F*p)/(F*p + n - p -1)
R2
```

---

**Problem 4**

You fit a multiple linear regression to a data of 20 observation and 4 predictors. You have determined that the coefficient of determination of the model is 0.8. Calculate the F-statistics to test the significant of the model.

(A) 4
(B) 8
(C) 12
(D) 15
(E) 18

```{r}
n = 20
p = 4
R2 = .8
F = (R2/p)/((1-R2)/(n-p-1))
F
```

---

### Generalized F-test

**Problem 5** (CAS Exam ST Spring 2016 Question 22)

The following two models were fit to 18 observations:

Model 1: $y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \epsilon$ \
Model 2: $y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_1x_2 +\beta_4 x_1^2 +\beta_5 x_2^2 + \epsilon$

The result of the regression are:

| Model Number | Error Sum of Squares | Regression Sum of Squares |
|:--------------|:----------------------|:---------------------------|
| 1            | 102                  | 23                        |
| 2            | 78                   | 47                        |

Calculate the value of the F-statistics used to test the hypothesis that $\beta_3 = \beta_4 =\beta_5 = 0$

(A) Less than 1.30
(B) At least 1.30, but less than 1.40
(C) At least 1.40, but less than 1.50
(D) At least 1.50, but less than 1.60
(E) At least 1.60


```{r}
n = 18
p = 5
q = 3

RSS0 = 102
RSS1 = 78
  
F = ((RSS0-RSS1)/q)/(RSS1/(n-p-1))
F
```

---

**Problem 6** 

A professor ran an experiment in three sections of a psychology courses to show that the more digits in a number, the more difficult it is to remember. The following variables were used in a multiple regression:

$x_1=$ number of digits in the number

$x_2 = 1$ if student was in section 1, 0 otherwise

$x_3 = 1$ if student was in section 2, 0 otherwise

$y=$ percentage of students correctly remembering the number

You are given

- A total of 42 students participated in the study
- The regression equation $$y = \beta_0 + \beta_1 x_1 + \beta_2 x_1^2 + \beta_3 x_2 +\beta_4 x_3 + \epsilon$$ was to fit the data and resulted in $R^2 = 0.940$
- A second regression equation $y = \gamma_0 + \gamma_1 x_1 + \gamma_2 x_1^2+\epsilon$ was to fit to the data and resulted $R^2 = 0.915$

Determine the F statistic used to test whether class section is significant variable. 


(A) 5.4
(B) 7.3
(C) 7.7
(D) 7.9
(E) 8.3

---

**Problem 7** 

You are determining the relationship of salary ($y$) to experience ($x_1$) for both men ($x_2 = 1$) and woman ($x_2=0$). You fit the model

$$y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_1x_2 + \epsilon$$

You are given

| Sources    | Sum of Squares | Degree of Freedom |
|:------------|:----------------|:-------------------|
| Regression | 330.0117       | 3                 |
| Error      | 12.8156        | 7                 |

You also fit the below model to the same observations. 

$$y = \gamma_0 + \gamma_1 x_1 +  \epsilon$$
You are given

| Sources    | Sum of Squares | Degree of Freedom |
|:------------|:----------------|:-------------------|
| Regression | 315.0992       | 1                 |
| Error      | 27.7281        | 9                 |

Determine the F-ratio to test whether the linear relationship between salary and experience is identical for men and women. 

(A) 0.6
(B) 2.0
(C) 3.5
(D) 4.1
(E) 6.2

---

<!--

**Problem 6** (SRM - Sample Question 19)

The regression model $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_1 X_2 + \epsilon$ is being investigated.
The following maximized log-likelihoods are obtained:

- Using only the intercept term: –1126.91
- Using only the intercept term, $X_1$, and $X_2$ : –1122.41
- Using all four terms: –1121.91

The null hypothesis $H_0 : \beta_1 = \beta_2 = \beta_3 = 0$ is being tested at the 5% significance level
using the likelihood ratio test.

Determine which of the following is true.

(A) The test statistic is equal to 1 and the hypothesis cannot be rejected.
(B) The test statistic is equal to 9 and the hypothesis cannot be rejected
(C) The test statistic is equal to 10 and the hypothesis cannot be rejected.
(D) The test statistic is equal to 9 and the hypothesis should be rejected.
(E) The test statistic is equal to 10 and the hypothesis should be rejected. 

-->

**Problem 8** (SRM - Sample Question 27)

Trevor is modeling monthly incurred dental claims. Trevor has 48 monthly claims
observations and three potential predictors:
- Number of weekdays in the month
- Number of weekend days in the month
- Average number of insured members during the month

Trevor obtained the following results from a linear regression:


|            | Coefficient    | Standard Error | Stat  | p-value |
|:------------|:----------------|:----------------|:-------|:---------|
| Intercept  | –45,765,767.76 | 20,441,816.55  | –2.24 | 0.0303  |
| Number of weekdays  | 513,280.76     | 233,143.23     | 2.20  | 0.0330  |
| Number of weekend days     |  280,148.46            | 483,001.55     | 0.58  | 0.5649  |
| Average number of members           | 38.64           | 6.42           | 6.01  | 0.0000  |

Determine which of the following variables should be dropped, using a 5% significance level.

I. Intercept \
II. Number of weekdays \
III. Number of weekend days \
IV. Number of members.

(A) I only
(B) II only
(C) III only
(D) IV only
(E) None should be dropped from the model

---

<!--
**Problem 9** (SRM - Sample Question 45)

Trisha runs a regression on a data set of n observations. She then calculates a 95%
confidence interval $(t,u)$ on y for a given set of predictors. She also calculates a
95% prediction interval $(v, w)$  on y for the same set of predictors.
Determine which of the following must be true. 

I. $lim_{n \to \infty}(u-t)$ \
II. $lim_{n \to \infty}(w-v)$ \
III. $w-v > u-t$ 


(A) None
(B) I and II only
(C) I and III only
(D) II and III only
(E) The correct answer is not given by (A), (B), (C), or (D). 

-->

