---
title: "AYU - Individual Week 4"
format: 
  html:
    toc: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
```

*Watch the videos to see how the similar problems are solve.*

## Random Walks

#### Problem 1 (SRM Sample - Question 4)

You are given:

i) The random walk model 

$$
y_t =  y_0 + c_1 + c_2 + c_3 +...+c_t,
$$

where $c_i, (i = 1, 2,..., t)$ denote observations from a white noise process. 

ii) The following ten observed values of $c_t$: 

| t     | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 |10|
|:-------|:----|:----|:----|:----|:----|:----|:----|:----|:----|:----|
| $y_t$ | 2  | 5  | 10  | 13  | 18  | 20  | 24  |25| 27  | 30  |


iii) $y_0 = 0$

Calculate the standard error of the 9 step-ahead forecast, $\hat{y}_{19}$ .

(A) 4/3
(B) 4
(C) 9
(D) 12
(E) 16 

```{r}

```


#### Problem 2 (SRM Sample - Question 55)

You are given the following eight observations from a time series that follows a
random walk model: 

| $t$     | 0  | 1  | 2  | 3  | 4   | 5   | 6   | 7   |
|:--------|:---|:---|:---|:---|:----|:----|:----|:----|
| $y_t$ | 3  | 5  | 7  | 8  | 12  | 15  | 21  | 22  |


You plan to fit this model to the first five observations and then evaluate it against the
last three observations using one-step forecast residuals. The estimated mean of the white
noise process is 2.25.

Let F be the mean error (ME) of the three predicted observations.

Let G be the mean square error (MSE) of the three predicted observations.

Calculate the absolute difference between F and G, $F-G$

(A) 3.48
(B) 4.31
(C) 5.54
(D) 6.47
(E) 7.63 


## Autoregressive

#### Problem 3 

A stationary autoregressive model of order one can be written as

$$y_t = \beta_0 + \beta_1y_{t-1} + \epsilon, t =  1, 2,...$$

Determine which of the following statements about this model is false

(A) The parameter $\beta_0$ must not equal 1.
(B) If the parameter $\beta_1 = 0$, then the model reduces to a white noise process.
(C) If the parameter $\beta_1 = 1$, then the model is a random walk.


#### Problem 4 (SRM Sample - Question 64)

You are given a stationary AR(1) model, 

$$
y_t = \beta_0 + \beta_1 y_{t-1} + \epsilon_t, t = 1, 2,..., T.
$$

Determine which or the following is always true.

(A) $\beta_0 \neq 0$
(B) $\beta_0 = 1$
(C) $\beta_1 = 0$
(D) $\beta_1 = 1$
(E) $|\beta_1| < 1$


#### Problem 5 (Parameter Estimation)

You are given the following six observed values of the autoregressive model of order one time series

$$
y_t = \beta_0 + \beta_1 y_{t-1} + \epsilon_t, \text{ with   }  Var(\epsilon_t) = \sigma^2.
$$

| $t$   | 1   | 2   | 3   | 4   | 
|:------|:----|:----|:----|:----|
| $y_t$ | 1   | 3   | 5   | 8   |

Calculate $\hat{\beta}_0$ and $\hat{\beta}_1$ using the conditional least squares method.

#### Problem 6 (SRM Sample - Question 58)

You are given the following six observed values of the autoregressive model of
order one time series 

$$
y_t = \beta_0 + \beta_1 y_{t-1} + \epsilon_t, \text{ with   }  Var(\epsilon_t) = \sigma^2.
$$

| $t$  | 1  | 2  | 3  | 4  | 5   | 6  |
|:-----|:---|:---|:---|:---|:----|:---|
| $y_t$| 31 | 35 | 37 | 41 | 45  | 51 |


The approximation to the conditional least squares method is used to estimate $\beta_0$
and $\beta_1$

Calculate the mean squared error $s^2$ that estimates $\sigma^2$

(A) 13
(B) 21
(C) 22
(D) 26
(E) 35 

#### Problem 7 (Forecasting with AR(1))

You are given an AR(1) with known parameters as follows.

\begin{align*}
y_t &= .6y_{t-1} -5 + \epsilon \\
y_T &= 7
\end{align*}

Calculate the three step ahead forecast of $y_{T+5}$


## Smoothing

#### Problem 8

You are given the following time series

| $t$   | 1   | 2   | 3   | 4   | 5   |
|:------|:----|:----|:----|:----|:----|
| $y_t$ | 1   | 3   | 5   | 8   | 13  |

Forecasting $y_{6}$ using smoothing average with $k=2$


#### Problem 9

You are given the following time series

| $t$   | 1   | 2   | 3   | 4   | 5   |
|:------|:----|:----|:----|:----|:----|
| $y_t$ | 1   | 3   | 5   | 8   | 13  |

Forecasting $y_{6}$ using exponential smoothing with $w = .8$

#### Problem 10 (SRM Sample - Question 46)

A time series was observed at times 0, 1, …, 100. The last four observations along
with estimates based on exponential and double exponential smoothing with $w =0.8$ are:

| Time ($t$)                  | 97   | 98   | 99   | 100   |
|:-----------------------------|:------|:------|:------|:-------|
| Observation ($y_t$)        | 96.9 | 98.1 | 99.0 | 100.2 |
| Estimates ($\hat{s}^(1)_t$) | 93.1 | 94.1 | 95.1 |       |
| Estimates ($\hat{s}^(2)_t$)                  | 88.9 | 89.9 |      |       |

All forecasts should be rounded to one decimal place and the trend should be rounded to
three decimal places.

Let F be the predicted value of $y_{102}$ using exponential smoothing with $w = 0.8$.

Let G be the predicted value of $y_{102}$ using double exponential smoothing with $w = 0.8$.

Calculate the absolute difference between F and G, $F-G$

(A) 0.0
(B) 2.1
(C) 4.2
(D) 6.3
(E) 8.4 
