---
title: "Week 4 - AYU - Individual"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
```

## Random Walks

#### Problem 1 (Sample - Question 4)

You are given:

i) The random walk model 

$$
y_t =  y_0 + c_1 + c_2 + c_3 +...+c_t,
$$

where $c_i, (i = 1, 2,..., t)$ denote observations from a white noise process. 

ii) The following ten observed values of $c_t$: 

| t     | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 |10|
|:-------|:----|:----|:----|:----|:----|:----|:----|:----|:----|:----|
| $y_t$ | 2  | 5  | 10  | 13  | 18  | 20  | 24  |25| 27  | 30  |


iii) $y_0 = 0$

Calculate the standard error of the 9 step-ahead forecast, $\hat{y}_{19}$ .

(A) 4/3
(B) 4
(C) 9
(D) 12
(E) 16 

```{r}

```


#### Problem 2 (Sample - Question 55)

You are given the following eight observations from a time series that follows a
random walk model: 

| $t$     | 0  | 1  | 2  | 3  | 4   | 5   | 6   | 7   |
|:--------|:---|:---|:---|:---|:----|:----|:----|:----|
| $y_t$ | 3  | 5  | 7  | 8  | 12  | 15  | 21  | 22  |


You plan to fit this model to the first five observations and then evaluate it against the
last three observations using one-step forecast residuals. The estimated mean of the white
noise process is 2.25.

Let F be the mean error (ME) of the three predicted observations.

Let G be the mean square error (MSE) of the three predicted observations.

Calculate the absolute difference between F and G, $F-G$

(A) 3.48
(B) 4.31
(C) 5.54
(D) 6.47
(E) 7.63 


#### Problem 3 (Sample - Question 21)


A random walk is expressed as

$$y_y = y_{t-1} + c_t$$

where

$E(c_t) = \mu_t$ and $Var(c_t) = \sigma_c^2$

Determine which statements is/are true with respect to a random walk model.

I. If $\mu_c \neq 0$, then the random walk is nonstationary in the mean. \
II. If $\sigma^2_c = 0$, then the random walk is nonstationary in the variance. \
III. If $\sigma^2_c > 0$, then the random walk is nonstationary in the variance.

(A) None
(B) I and II only
(C) I and III only
(D) II and III only
(E) The correct answer is not given by (A), (B), (C), or (D). 

#### Problem 4 (Sample - Question 31)

Determine which of the following indicates that a nonstationary time series can be
represented as a random walk

I. A control chart of the series detects a linear trend in time and increasing
variability.

II. The differenced series follows a white noise model.

III. The standard deviation of the original series is greater than the standard
deviation of the differenced series.

(A) I only
(B) II only
(C) III only
(D) I, II and III
(E) The correct answer is not given by (A), (B), (C), or (D).

## Autoregressive

#### Problem 5 (Sample - Question 22)

A stationary autoregressive model of order one can be written as

$$y_t = \beta_0 + \beta_1y_{t-1} + \epsilon, t =  1, 2,...$$

Determine which of the following statements about this model is false

(A) The parameter $\beta_0$ must not equal 1.
(B) The absolute value of the parameter $\beta_1$ must be less than 1.
(C) If the parameter $\beta_1 = 0$, then the model reduces to a white noise process.
(D) If the parameter $\beta_1 = 1$, then the model is a random walk.
(E) Only the immediate past value, $y_{t-1}$, is used as a predictor for $y_t$ . 


#### Problem 6 (Sample - Question 58)

You are given the following six observed values of the autoregressive model of
order one time series 

$$
y_t = \beta_0 + \beta_1 y_{t-1} + \epsilon_t, \text{ with   }  Var(\epsilon_t) = \sigma^2.
$$

| $t$  | 1  | 2  | 3  | 4  | 5   | 6  |
|:-----|:---|:---|:---|:---|:----|:---|
| $y_t$| 31 | 35 | 37 | 41 | 45  | 51 |


The approximation to the conditional least squares method is used to estimate $\beta_0$
and $\beta_1$

Calculate the mean squared error $s^2$ that estimates $\sigma^2$

(A) 13
(B) 21
(C) 22
(D) 26
(E) 35 

#### Problem 7 (Sample - Question 64)

You are given a stationary AR(1) model, 

$$
y_t = \beta_0 + \beta_1 y_{t-1} + \epsilon_t, t = 1, 2,..., T.
$$

Determine which or the following is always true.

(A) $\beta_0 \neq 0$
(B) $\beta_0 = 1$
(C) $\beta_1 = 0$
(D) $\beta_1 = 1$
(E) $|\beta_1| < 1$

#### Problem 8

You are given
\begin{align*}
y_t &= .6y_{t-1} -5 + \epsilon \\
y_T &= 7
\end{align*}
Calculate the three step ahead forecast of $y_{T+5}$


## Smoothing

#### Problem 9 (Sample - Question 46)

A time series was observed at times 0, 1, …, 100. The last four observations along
with estimates based on exponential and double exponential smoothing with $w =0.8$ are:

| Time ($t$)                  | 97   | 98   | 99   | 100   |
|:-----------------------------|:------|:------|:------|:-------|
| Observation ($y_t$)        | 96.9 | 98.1 | 99.0 | 100.2 |
| Estimates ($\hat{s}^(1)_t$) | 93.1 | 94.1 | 95.1 |       |
| Estimates ($\hat{s}^(2)_t$)                  | 88.9 | 89.9 |      |       |

All forecasts should be rounded to one decimal place and the trend should be rounded to
three decimal places.

Let F be the predicted value of $y_{102}$ using exponential smoothing with $w = 0.8$.

Let G be the predicted value of $y_{102}$ using double exponential smoothing with $w = 0.8$.

Calculate the absolute difference between F and G, $F-G$

(A) 0.0
(B) 2.1
(C) 4.2
(D) 6.3
(E) 8.4 


<!---


#### Problem (Sample - Question 38)

You are given two models:

Model L: 
 
$$y_t = \beta_0 + \beta_1 t + \epsilon_t$$
where $\epsilon_t$ is a white noise process, for $t = 0,1,2, …$

Model M: 
\begin{align*}
y_t &= y_0 + \mu_c t + u_t \\
c_t &= y_t - y_{t-1} \\
u_t &= \sum_{j=1}^t \epsilon_j
\end{align*}

where $\epsilon_t$ is a white noise process, for $t = 0,1,2, …$
Determine which of the following statements is/are true.

I. Model L is a linear trend in time model where the error component is not a
random walk.

II. Model M is a random walk model where the error component of the model
is also a random walk.

III. The comparison between Model L and Model M is not clear when the
parameter $\mu_c = 0$

(A) I only
(B) II only
(C) III only
(D) I, II and III
(E) The correct answer is not given by (A), (B), (C), or (D). 

-->

<!---

#### Problem (Sample - Question 3)

You are given:

i) The random walk model 

$$
y_t =  y_0 + c_1 + c_2 + c_3 +...+c_t,
$$

where $c_i, (i = 1, 2,..., t)$ denote observations from a white noise process. 

ii) The following nine observed values of $c_t$: 

| t     | 11 | 12 | 13 | 14 | 15 | 16 | 17 | 18 | 19 |
|:-------|:----|:----|:----|:----|:----|:----|:----|:----|:----|
| $c_t$ | 2  | 3  | 5  | 3  | 4  | 2  | 4  | 1  | 2  |

iii) The average value of $c_1, c_2,..., c_{10}$ is 2. 

iv) The 9 step ahead forecast of $y_{19}$,  $\hat{y}_{19}$ is estimated based on the observed value
of 10 y . 

Calculate the forecast error,  $y_{19} - \hat{y}_{19}$

(A) 1
(B) 2
(C) 3
(D) 8
(E) 18 

-->


