---
title: "Week 1 - AYU - Individual"

output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=FALSE, warning = FALSE, message = FALSE)
```

### Parameter estimation

**Problem 1.** Similar Problems: Example 1.1.2

| x | 2 | 3 | 5 | 6 | 1 | 9 |10| 15|
|:---|:---|:---|:---|:---|:---|:---|:---|:---|
| y | 1 | 4 | 6 | 4 | 4 | 3 |20|25|

Write the equation of the best fitted line.

A. $y = -1.722 + 1.584x$ \
B. $y = -1.722 - 1.584x$ \
C. $y = 1.722 + 1.584x$ \
D. $y = 1.584 + 1.722x$ \
E. $y = -1.584 + 1.722x$ \

```{r, echo = FALSE, eval=FALSE}
x = c(2,	3,	5,	6,	1,	9, 10, 15)
y = c(1,	4,	6,	4,	4,	3, 20, 25)
lm(y~x)
```


**Problem 2.** (Similar Problem: Example 1.1.3)

The regression model is $y=\beta_0 + \beta_1x+\epsilon$. There are six observations. The summary statistics are: 
\begin{align*} 
\sum{y_i} &= 58, \\
\sum{x_i}&=21, \\
\sum x_i^2 &= 91, \\
\sum x_iy_i &= 259, \\
\sum y_i^2 &= 754
\end{align*}

Calculate the least squares estimate of $\beta_1$.

(A) 3.0
(B) 3.2
(C) 3.4
(D) 3.6
(E) 3.8

```{r}
x = c(1,2,3,4,5,6)
y = c(2,6,7,11, 12, 20)
n= length(x)
beta_1 = (sum(x*y)-n*mean(x)*mean(y))/(sum(x^2)-n*(mean(x))^2)

l = lm(y~x)
beta = l$coefficients[2]
```


**Problem 3** (Similar Problem: Example 1.1.3)

The regression model is $y=\beta_0 + \beta_1x+\epsilon$. You are given the follows. 
\begin{align*} 
n &= 10, \\
\bar{y} &= 21.1, \\
\bar{x}&=7.5, \\
\sum x_i^2 &= 759,\\ 
\sum x_iy_i &= 2253,\\ 
\sum y_i^2 &= 7657
\end{align*}

```{r, echo=FALSE}
x = c(1,2,5,7,9, 10, 3, 11, 12, 15)
y = c(3, 6, 10, 11,15,20, 15, 21, 50, 60)
```

Predict $y_{11}$ given that $x_{11}=20$

A. 63.748 \
B. 62.758 \
C. 64.758 \
D. 65.758 \
E. The correct answer is not given by (A), (B), (C), or (D).

**Problem 4** (Similar Problem: Example 1.1.4)

You are given the following summary statistics:

\begin{align*}
\bar{x} = 6 \\
\bar{y}=13.75 \\
\sum (x_i-\bar{x})^2 = 102 \\
\sum (x_i - \bar{x})(y_i-\bar{y}) = 192 \\
\sum (y_i-\bar{y})^2 = 503.5
\end{align*}

Determine the equation of the regression line, using the least squares method. 

(A) $y = 2.456  + 1.882x$ 

(B) $y =  0.78  + 1.882x$

(C) $y =  2.456  + 0.65x$

(D) $y =  0.39  + 0.70x$

(E) The correct answer is not given by (A), (B), (C), or (D).

```{r, echo=FALSE}
x = c(1,2,5,7,9, 10, 3, 11)
y = c(3, 6, 10, 11,15,20, 15, 30)

beta_1 = sum((x-mean(x))*(y-mean(y)))/sum((x-mean(x))^2)
lm(y~x)
```

### Goodness of Fit

**Problem 5** (Similar Problem: Example 1.2.1)

For a simple linear regression model the sum of squares of the residuals is 

$$
\sum_{i=1}^{25} e_i^2 = 300
$$
and the $R^2$ statistic is 0.8. Calculate the total sum of squares (TSS) for this model. 

(A) 1000
(B) 1200
(C) 1500
(D) 2000
(E) 2500

**Problem 6** (Similar Problem: Example 1.2.1)

For a simple linear regression model the total sum of squares (TSS) is 1000

and the $R^2$ statistic is 0.7. Calculate the sum of squares of the residuals for this model. 

(A) 300
(B) 400
(C) 500
(D) 600
(E) None of the above


**Problem 7** (Similar Problem: Example 4.1.3 or SRM - Sample Question 11)

You are given the following results from a regression model.

| Observation number (i) | $y_i$ | $\hat{f}(x_i)$ |
|:------------------------|:-------|:----------------|
| 1                      | 1     | 4              |
| 2                      | 2     | 3              |
| 3                      | 6     | 7              |
| 4                      | 8     | 9              |
| 5                      | 4     | 6              |

Calculate the sum of squared errors (SSE).

(A) 10
(B) 12
(C) 14
(D) 16
(E) 46

```{r}
y = c(1,2,6,8,4)
y_hat = c(4,3,7,9,6)

SSE = sum((y-y_hat)^2)

```

---

**Problem 8** (Similar Problem Example 1.2.5 or SRM - Sample Question 44)

Two actuaries are analyzing dental claims for a group of $n = 100$ participants. The predictor variable is sex, with 0 and 1 as possible values. 

Actuary 1 uses the following regression model: 

$$Y = \beta + \epsilon$$
Actuary 2 uses the following regression model: 

$$Y = \beta_0 + \beta_1 \times Sex + \epsilon$$

The residual sum of squares for the regression of Actuary 2 is 100,000 and the total sum
of squares is 120,000.

Calculate the F-statistic to test whether the model of Actuary 2 is a significant improvement over the model of Actuary 1. 

(A) 20.5
(B) 22.6
(C) 19.6
(D) 30.1
(E) 34.5

```{r}
n = 100
RSS = 100000
TSS = 120000
Reg_SS = TSS- RSS
F = Reg_SS/(RSS/(n-2))
```


**Problem 9** (Similar Problem Example 1.2.5 or SRM - Sample Question 44)

Two actuaries are analyzing dental claims for a group of $n = 100$ participants. The predictor variable is sex, with 0 and 1 as possible values. 

Actuary 1 uses the following regression model: 

$$Y = \beta + \epsilon$$

Actuary 2 uses the following regression model: 

$$Y = \beta_0 + \beta_1 \times Sex + \epsilon$$

Given $R^2 = .7$. Calculate the F-statistic to test whether the model of Actuary 2 is a significant
improvement over the model of Actuary 1. 

(A) 120
(B) 135
(C) 147
(D) 157
(E) 240

```{r}
n = 100
R2 = .6
F = (n-2)*(R2/(1-R2))
```



**Problem 10** (t-test)

The results of fitting ten observation by the regression model, $y=\beta_0 + \beta_1x+\epsilon$, are given below.

Determine the test results of the hypothesis $H_0: \beta_1 = 0$ against $H_{\alpha}: \beta_1 \neq 0$.

|           | Estimate | Std. Error | t value | Pr(>\|t\|) |
|:-----------|:----------|:------------|:---------|:------------|
| Intercept | -4.4916  | 6.6540     | -0.675  | 0.51869    |
| x         | 3.4122   | 0.7638     | 4.468   | 0.00209    |

A. Reject at $\alpha = .2$ \
B. Reject at $\alpha = .1$ \
C. Reject at $\alpha = .05$ \
D. Reject at $\alpha = .01$ \
E. All (A), (B), (C), or (D) are correct.




### Application of Linear Model.

**Problem 11** (SRM - Sample Question 23)

Toby observes the following coffee prices in his company cafeteria:

- 12 ounces for 1.00
- 16 ounces for 1.20
- 20 ounces for 1.40

The cafeteria announces that they will begin to sell any amount of coffee for a price that
is the value predicted by a simple linear regression using least squares of the current
prices on size.

Toby and his co-worker Karen want to determine how much they would save each day,
using the new pricing, if, instead of each buying a 24-ounce coffee, they bought a 48-
ounce coffee and shared it.

Calculate the amount they would save.

(A) It would cost them 0.40 more.
(B) It would cost the same.
(C) They would save 0.40.
(D) They would save 0.80.
(E) They would save 1.20. 


**Problem 12** 

Peter observes the following coffee prices in his company cafeteria:

- 1 bagel for 1.00 (USD)
- 2 bagel for 1.50 (USD)

The cafeteria announces that they will begin to sell any amount of bagels for a price that
is the value predicted by a simple linear regression using least squares of the current
prices.

With the new pricing model, how much Peter would save if he bought 10 bagels instead of 5 bagels twice?

(A) It would cost him more so he would not save any money. 
(B) It would cost the same.
(C) He would save 0.5 (USD)
(D) He would save 1 (USD)
(E) He would save 1.5 (USD)


<!--

**Problem 10** (Calculating t-statistics given a summarized output) Similar Problem: Problem 1.3.5

Calculate the t-value of a linear regression model, $y=\beta_0 + \beta_1x+\epsilon$, given that

\begin{align*}
\sum_{i=1}^{10}(x_i-\hat{x})(y_i-\hat{y}) = 400\\
\sum_{i=1}^{10} x_i = 50 \\
\sum_{i=1}^{10} x_i^2 = 650 \\
s^2 = 169 
\end{align*}

A. 1.538 \
B. 2.538 \
C. 0.285 \
D. 0.101 \
E. 1.132

-->

<!-- 


**Problem 12** (F-test)

The results of fitting ten observation by the regression model, $y=\beta_0 + \beta_1x+\epsilon$, are given below.

Determine the p-value of the F-test for $H_0: \beta_1 = 0$ against $H_{\alpha}: \beta_1 \neq 0$.

|           | Estimate | Std. Error | t value | Pr(>\|t\|) |
|:-----------|:----------|:------------|:---------|:------------|
| Intercept | 3.30000      | 1.39766     | 2.361     | 0.0776     |
| x         | 0.08462       | 0.27410        | 0.309      | 0.7730      |


A. p-value of the F-test is not given in the result \
B. p-value = 0.0776 \
C. p-value = 0.7730 \
D. p-value = $0.7730^2 = 0.597529$ \
E. The correct answer is not given by (A), (B), (C), or (D).


---

### Type 3: Statistical Inference: Confidence Intervals and prediction Interval

**Problem 12** (SRM - Sample Question 13)

In a simple linear regression model based on over 100 observations, you are given
the following estimates.

i) The estimated slope is –1.03.
ii) The standard error of the estimated slope is 0.06.

Calculate the 95% confidence interval for the slope.

(A) (–1.15, –0.91)
(B) (–1.13, –0.93)
(C) (–1.11, –0.95)
(D) (–1.09, –0.97)
(E) (–1.07, –0.99)


**Problem 13** (SRM - Sample Question 13)
Determine which of the following statements is/are true for a simple linear
relationship, $y = \beta_0 + \beta_1 x + \epsilon$

- I. If $\epsilon = 0$ , the 95% confidence interval is equal to the 95% prediction
interval
- II. The prediction interval is always at least as wide as the confidence
interval. 
- III. The prediction interval quantifies the possible range for $E(y|x)$

(A) I only
(B) II only
(C) III only
(D) I, II, and III
(E) The correct answer is not given by (A), (B), (C), or (D)



**Problem 14** (SRM - Sample Question 56)

Determine which of the following statements about prediction is true.

(A) Each of several candidate regression models must produce the same prediction.
(B) When making predictions, it is assumed that the new observation follows the
same model as the one used in the sample.
(C) A point prediction is more reliable than an interval prediction.
(D) A wider prediction interval is more informative than a narrower prediction
interval.
(E) A prediction interval should not contain the single point prediction

-->