---
title: "Simple Linear Model"
format: beamer
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Simple Linear Model

-   Given the data

| $x$ | $y$ |
|-----|-----|
| 1   | 1.1 |
| 2   | 1.8 |
| 3   | 2.7 |
| 4   | 4.5 |

# Scatter plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)

x = c(1,2,3,4)
y = c(1.1, 1.8, 2.7, 4.5)

d1 = as_tibble(data.frame(
     x = x, 
     y = y))

# functions
plot_line <- function(x1, y1, x2, y2)
  
{
  slope = (y2-y1)/(x2-x1)
  intercept =  y1 - slope*x1
  f = function(xx){return(xx*slope + intercept)}
  
  pp = ggplot(data = d1, aes(x = x, y=y))+
  geom_point()+
    geom_segment(aes(x = 0.5, y = f(0.5), xend = 4.5, yend = f(4.5)))+
    labs(title = paste("Distance:"),
         subtitle = paste0("y = ", round(slope,2), "x+", round(intercept,2)))+
    xlim(0,5)+
    ylim(0,5)+
    coord_fixed()

  print(pp)

}

plot_line2 <- function(x1, y1, x2, y2)
{
slope = (y2-y1)/(x2-x1)
intercept =  y1 - slope*x1

f = function(xx){return(xx*slope + intercept)}

squared_errors = sum((y-f(x))^2)

pp = ggplot(data = d1, aes(x = x, y=y))+
  geom_point()+
  geom_segment(aes(x = 0.5, y = f(0.5), xend = 4.5, yend = f(4.5)))+
  geom_segment(aes(x = x[1], y = y[1], xend = x[1], yend = f(x[1])), color = 'red')+
  geom_segment(aes(x = x[2], y = y[2], xend = x[2], yend = f(x[2])), color = 'red')+
  geom_segment(aes(x = x[3], y = y[3], xend = x[3], yend = f(x[3])), color = 'red')+
  geom_segment(aes(x = x[4], y = y[4], xend = x[4], yend = f(x[4])), color = 'red')+
  geom_text(x = x[1], y = .5*(y[1]+f(y[1])), label = paste(round(abs(y[1]-f(x[1])),2)))+
  geom_text(x = x[2], y = .5*(y[2]+f(y[2])), label = paste(round(abs(y[2]-f(x[2])),2)))+
  geom_text(x = x[3], y = .5*(y[3]+f(y[3])), label = paste(round(abs(y[3]-f(x[3])),2)))+
  geom_text(x = x[4], y = .5*(y[4]+f(y[4])), label = paste(round(abs(y[4]-f(x[4])),2)))+
  labs(title = paste("Distance:", round(squared_errors,2)),
         subtitle = paste0("y = ", round(slope,2), "x+", round(intercept,2)))+
  xlim(0,5)+
  ylim(0,5)+
  coord_fixed()

print(pp)

}

plot_line3 <- function()
  
{
  l = lm(y~x) 
slope = l$coefficients[2]
intercept =  l$coefficients[1]
f = function(xx){return(xx*slope + intercept)}
squared_errors = sum((y-f(x))^2)
pp = ggplot(data = d1, aes(x = x, y=y))+
  geom_point()+
  geom_segment(aes(x = 0.5, y = f(0.5), xend = 4.5, yend = f(4.5)))+
  geom_segment(aes(x = x[1], y = y[1], xend = x[1], yend = f(x[1])), color = 'red')+
  geom_segment(aes(x = x[2], y = y[2], xend = x[2], yend = f(x[2])), color = 'red')+
  geom_segment(aes(x = x[3], y = y[3], xend = x[3], yend = f(x[3])), color = 'red')+
  geom_segment(aes(x = x[4], y = y[4], xend = x[4], yend = f(x[4])), color = 'red')+
  geom_text(x = x[1], y = .5*(y[1]+f(y[1])), label = paste(round(abs(y[1]-f(x[1])),2)))+
  geom_text(x = x[2], y = .5*(y[2]+f(y[2])), label = paste(round(abs(y[2]-f(x[2])),2)))+
  geom_text(x = x[3], y = .5*(y[3]+f(y[3])), label = paste(round(abs(y[3]-f(x[3])),2)))+
  geom_text(x = x[4], y = .5*(y[4]+f(y[4])), label = paste(round(abs(y[4]-f(x[4])),2)))+
  labs(title = paste("Distance:", round(squared_errors,2)),
         subtitle = paste0("y = ", round(slope,2), "x+", round(intercept,2)))+
  xlim(0,5)+
  ylim(0,5)+
  coord_fixed()

print(pp)

}

```

```{r}
# ploting the point
ggplot(data = d1, aes(x = x, y=y))+
  geom_point()+
    xlim(0,5)+
    ylim(0,5)+
    coord_fixed()
```

# Which line is closer to the points?

```{r}
x1 = 0
y1 = 1
x2 = 5
y2 = 3

a1 = 0.5
b1 = 0
a2 = 5
b2 = 4

ggplot(data = d1, aes(x = x, y=y))+
  geom_point()+
  geom_segment(aes(x = a1, y = b1, xend = a2, yend = b2))+
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2, color = 'red'))+
  xlim(0,5)+
  ylim(0,5)+
  coord_fixed()+
  theme(legend.position="none")

```

# Squared Distance between a line and points

```{r}
plot_line(x1, y1, x2, y2)
```

# Squared Distance between a line and points

```{r}
plot_line2(x1, y1, x2, y2)
```

# Squared Distance between a line and points

```{r}
plot_line(a1, b1, a2, b2)
```

# Squared Distance between a line and points

```{r}
plot_line2(a1, b1, a2, b2)
```

# What is the closest line to the points?

-   The least squared methods give us the formula for the `closest` line:

-   $y = \hat{\beta_1}x+\hat{\beta_0}$

-   $\hat{\beta_{1}} = \frac{\sum_{i=1}^{n}{(x_i-\bar{x})(y_i-\bar{y})}}{\sum_{i=1}^{n}(x_i-\bar{x})^2} = \frac{S_{xy}}{S_{xx}}$

-   $\hat{\beta_{0}} = \bar{y} - \hat{\beta_{1}}\bar{x}$

-   This line is also called the best fitted line

# Calculation

| $x$ | $y$ |
|-----|-----|
| 1   | 1.1 |
| 2   | 1.8 |
| 3   | 2.7 |
| 4   | 4.5 |

# Calculation

|        | $x$ | $y$ | $xy$ | $x^2$ |
|:------:|:---:|:---:|:----:|:-----:|
|        |  1  | 1.1 |      |       |
|        |  2  | 1.8 |      |       |
|        |  3  | 2.7 |      |       |
|        |  4  | 4.5 |      |       |
| $\sum$ |     |     |      |       |

# Calculation

|        | $x$ | $y$ | $xy$ | $x^2$ |
|:------:|:---:|:---:|:----:|:-----:|
|        |  1  | 1.1 |      |       |
|        |  2  | 1.8 |      |       |
|        |  3  | 2.7 |      |       |
|        |  4  | 4.5 |      |       |
| $\sum$ |     |     |      |       |

-   $\bar{x} = \frac{1+2+3+4}{4} = 2.5$
-   $\bar{y} = \frac{1.1+1.8+2.4+4.5}{4} = 2.525$

# Calculation

|        | $x$ | $y$ | $xy$ | $x^2$ |
|:------:|:---:|:---:|:----:|:-----:|
|        |  1  | 1.1 | 1.1  |   1   |
|        |  2  | 1.8 | 3.6  |   4   |
|        |  3  | 2.7 | 8.1  |   9   |
|        |  4  | 4.5 |  18  |  16   |
| $\sum$ |     |     | 30.8 |  30   |

-   $\hat{\beta_1} = \frac{\sum xy -n\bar{x}\bar{y}}{\sum x^2 - n\bar{x}^2} = 1.11$

-   $\hat{\beta_0} = \bar{y} - \hat{\beta_{1}}\bar{x} = -0.25$

# Graph

```{r}
plot_line3()
```

# Some other lines

```{r}
plot_line2(0, 0, 5, 5)
```

# Some other lines

```{r}
plot_line2(0, 1, 5, 4)
```

# Some other lines

```{r}
plot_line2(0, 2, 5, 3)
```

# Some other lines

```{r}
plot_line2(0, 3, 5, 2)
```

# Sum Up

-   The best fitted line or the least squared line is the line that is closet to the data point in term of the squared distance.
