
---
title: "Week 1 - Pod - Submission"
output:
  word_document: default
  pdf_document: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

## Question 1

Given the following data

| x   | 2   | 3   | 5   | 6   | 1   | 9   | 10  | 15  |
|:----|:----|:----|:----|:----|:----|:----|:----|:----|
| y   | 1   | 4   | 6   | 4   | 4   | 3   | 20  | 25  |

a.  Calculate the mean, standard deviation, and variance of $x$
b.  Calculate the correlation of $x$ and $y$
c.  Run a linear model when regressing $y$ on $x$.

```{r}

```


------------------------------------------------------------------------

## Question 2*

Given the following data

| x   | 2   | 3   | 5   | 6   | 1   |
|:----|:----|:----|:----|:----|:----|
| y   | 1   | 4   | 6   | 4   | 4   |

a.  Calculate the mean, standard deviation, and variance of $x$
b.  Calculate the correlation of $x$ and $y$
c.  Run a linear model when regressing $y$ on $x$.

```{r}

```

------------------------------------------------------------------------

## Question 3

a.  Import in R the `Automobile UK Collision Claims` dataset at

-   [Collision Claims](https://instruction.bus.wisc.edu/jfrees/jfreesbooks/Regression%20Modeling/BookWebDec2010/CSVData/AutoCollision.csv) or [Alternative Link](data/AutoCollision.csv)

b.  Use the View function to view the data


```{r}

```

------------------------------------------------------------------------

## Question 4

Run linear regression to regress the `Claim_Count` on the `Severity`.  Give the summary of the model. 

```{r}

```

------------------------------------------------------------------------

## Question 5*

In this question, we will use the `Automobile Insurance Claims` dataset at the below link:

[Automobile Insurance Claims](https://instruction.bus.wisc.edu/jfrees/jfreesbooks/Regression%20Modeling/BookWebDec2010/CSVData/NAICExpense.csv)

a. Import the following dataset into Rstudio. 
b. View the dataset. 
c. Run linear model to regress `PAID` on `AGE`. Report the summary of this model. 

```{r}

```

------------------------------------------------------------------------
